---
title: "Case Study 2"
author: "Donald Anderson"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(caret)
library(Metrics)
library(rpart)
library(readxl)
```

## Challenge

## Talent management is defined as the iterative process of developing and retaining employees. It may include workforce planning, employee training 
## programs, identifying high-potential employees and reducing/preventing voluntary employee turnover (attrition). In order to gain a competitive edge 
## over its competition, the CEO and CFO of Frito Lay is planning to leverage data science for talent management. Frito Lays’ leadership team has determined that 
## predicting employee turnover will be its first application of data science for talent management.

## Solution

## DDS Analytics specializes in talent management solutions for Fortune 100 companies. Their client, Frito Lay has provided the analytics company with a 
## dataset to conduct data analysis that will identify factors that lead to attrition. DDS Analytics will present the executive staff with evidence-based 
## analysis of the top three factors that contribute to employee turnover. Job-role specific trends that exist in the dataset will be shared in their 
## findings as well as other interesting trends and observations from their research.

```{r}
data <- read.csv(file.choose(), header = TRUE)
head(data)
```

## Exploratory Data Analysis:

## Box Plots of Attrition vs Monthly Income, Years at Company, & Age
```{r}
par(mfrow=c(1,1))
boxplot(data$MonthlyIncome ~ data$Attrition, 
        main = "Monthly Income vs Attrition Status",
        col = "purple",
        xlab = "Attrition",
        ylab = "Monthly Income")

boxplot(data$YearsAtCompany ~ data$Attrition, 
        main = "Years At Company vs Attrition Status",
        col = "red",
        xlab = "Attrition",
        ylab = "Years At Company")

boxplot(data$Age ~ data$Attrition,
        main = "Age vs Attrition Status",
        col = "steel blue",
        xlab = "Attrition",
        ylab = "Age")
```

## Pie Chart of Attrition by Department
```{r}
data_1 <- data %>%
  filter(Attrition == "Yes")
temp <- as.data.frame(table(data_1$Department))
colnames(temp) <- c("Department", "Count")
temp <- temp %>%
  mutate(Proportion = round((Count / sum(Count)) * 100, 2))

ggplot(temp, 
       aes(x = "", 
           y = Proportion, 
           fill = Department)) +
  geom_col() +
  geom_text(aes(label = round(Proportion, 2)),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  labs(title = "Department Wide Attrition Percentage")
```

## Barplot by Job Role
```{r}
temp_1 <- as.data.frame(table(data_1$JobRole))
colnames(temp_1) <- c("Job Role", "Count")
temp_1 <- temp_1 %>%
  mutate(Proportion = round((Count / sum(Count)) * 100, 2))

ggplot(data = temp_1, 
       mapping = aes(x = `Job Role`, 
                     y = Proportion, 
                     fill = `Job Role`)) +
  geom_bar(stat="identity", 
           position = "dodge") +
  theme(axis.text.x = element_text(angle=90, hjust=1)) +
  labs(title = "Attrition's Percentage By Job Role")
```
## Pie Chart of Attrition by Avg. Distance from Home
```{r}
temp_2 <- data %>%
  group_by(Attrition) %>%
  summarise(Average_Distance = mean(DistanceFromHome, na.rm = T))
temp_2$Average_Distance <- round(temp_2$Average_Distance , 2)

ggplot(temp_2, 
       aes(x = "", 
           y = Average_Distance, 
           fill = Attrition)) +
  geom_col() +
  geom_text(aes(label = round(Average_Distance, 2)),
            position = position_stack(vjust = 0.5)) +
  coord_polar(theta = "y") +
  labs(title = "Attrition Status By Avg Distance from Home")
```

## Predictive Model

## Data Preparation:

```{r}
colSums(is.na(data))
```
```{r}
data <- data[, 2:36]
str(data)
```

```{r}
data$Over18 <- as.factor(data$Over18)
data$OverTime <- as.factor(data$OverTime)
data$MaritalStatus <- as.factor(data$MaritalStatus)
data$JobRole <- as.factor(data$JobRole)
data$Gender <- as.factor(data$Gender)
data$EducationField <- as.factor(data$EducationField)
data$Department <- as.factor(data$Department)
data$BusinessTravel <- as.factor(data$BusinessTravel)
data$Attrition <- as.factor(data$Attrition)
```

```{r}
lapply(data, n_distinct)
```

```{r}
data <- data[, -which(names(data) %in% c("Over18", 
                                         "StandardHours"))]
```
## Data Partition:

```{r}
set.seed(123)
split1<- sample(c(rep(0, 0.7 * nrow(data)), 
                  rep(1, 0.3 * nrow(data))))
train <- data[split1 == 0, ] 
test <- data[split1== 1, ]
```
## Decision Tree Classifier:
```{r}
library(rpart)
fit_DT <- rpart(Attrition ~.,
                data = train,
                method = "class")
summary(fit_DT)
```
## Evaluate On Test Set & Provided Test Set:

## Perform prediction on test set from data splitting step
```{r}
pred <- predict(fit_DT, 
                test,
                type = "class")
cm <- table(pred, test$Attrition)
confusionMatrix(cm)
```
# Load attrition data without labels.
```{r}
data_attrition <- read.csv(file.choose(), header = TRUE)
```

## Predict the Attrition and write file as csv
```{r}
pred <- predict(fit_DT, 
                data_attrition,
                type = "class")
data_attrition_res <- as.data.frame(pred)
data_attrition_res$ID <- data_attrition$ID
head(data_attrition_res)

write.csv(data_attrition_res, "Case2PredictionsAnderson Attrition.csv", row.names = FALSE)
```

## Regression for Salary

## Build Regression Model:
```{r}
model_regression <- lm(MonthlyIncome ~., 
                       data = train)
summary(model_regression)
```

## Perform predictions and calculate RMSE
```{r}
pred <- predict(model_regression, 
                test)
rmse(pred, 
     test$MonthlyIncome)
```


# Read in data from Excel file
```{r}
data_test <- read_excel(file.choose())

# Perform prediction and save results as csv.
pred <- predict(model_regression, 
                data_test)

salary_res <- as.data.frame(pred)
salary_res$ID <- data_test$ID
head(salary_res)

write.csv(salary_res, "Case2PredictionsAnderson Salary.csv", row.names = FALSE)
```







